Efficient High-Level Abstractions for Web Programming

### 2013-09-11 papers/gpce2013/gpce2013.tex
abstract reworded

Writing large Web applications is known to be difficult. One challenge comes from the fact that the
application's logic is scattered into heterogeneous clients and servers, making it difficult to
share code between both sides or to move code from one side to the other. Another challenge is
performance: while Web applications rely on ever more code on the client-side, they may run on smart
phones with limited hardware capabilities. These two challenges raise the following problem: how to
benefit from high-level languages and libraries making code complexity easier to manage and
abstracting over the clients and servers differences without trading this ease of engineering for
performance? This article presents high-level abstractions defined as deep embedded DSLs in Scala,
that can (1) generate efficient code leveraging the target platform characteristics, (2) be shared
between clients and servers. Though code written with our DSL has a high level of abstraction, our
benchmark on a real world application reports that it runs as fast as hand tuned low-level
JavaScript code.
\end{abstract}

>>>>

Writing large Web applications is known to be difficult. One challenge comes from the fact that the
application's logic is scattered into heterogeneous clients and servers, making it difficult to
share code between both sides or to move code from one side to the other. Another challenge is
performance: while Web applications rely on ever more code on the client-side, they may run on smart
phones with limited hardware capabilities. These two challenges raise the following problem: how to
benefit from high-level languages and libraries making code complexity easier to manage and
abstracting over the clients and servers differences without trading this ease of engineering for
performance? This article presents high-level abstractions defined as deep embedded DSLs in Scala that can generate efficient code leveraging the characteristics of both client and server environments. We compare performance on client-side against other candidate technologies and against hand written low-level JavaScript code. Though code written with our DSL has a high level of abstraction, our
benchmark on a real world application reports that it runs as fast as hand tuned low-level
JavaScript code.
\end{abstract}
